---
title: 'Lab #5'
author: "Jung Ho Suh"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
per <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)

```


***1. Create a data frame containing the persons who are fatally hurt in the accidents (see FARS manual and look up variable INJ_SEV)***

```{r}
library(dplyr)
library(tidyr)

fatal_hurt <- per %>% 
  filter(INJ_SEV == 4)

```

***2. Create a data frame containing the most dangerous vehicle make in each state. The number of persons fatally hit in the vehicle make is used to assess the (non-)safety of a make. Make sure to handle the missing values appropriately. (look up variable MAKE)***

```{r}
library(dplyr)
library(tidyr)

make_by_fatal_hit <- fatal_hurt %>%
  group_by(MAKE,STATE) %>%
  mutate(
    total = n()
  ) %>%
  filter(!(is.na(MAKE))) %>%
  select(c(STATE,MAKE,total)) %>%
  arrange(STATE,desc(total)) %>%
  distinct() %>%
  ungroup() %>%
  group_by(STATE) %>%
  slice(1)

```


***3. Create a map, and label each state with the most dangerous vehicle. Discuss the definition of the most dangerous vehicle, and what you find from the map. (Hint: Read the description for the STATE and COUNTY columns in the FARS manual. The state & county codes are Geographic Locator Codes (GLCs) from the General Services Administration’s (GSA) publication. Use readxl::read_xlsx to read in the GLCs.)***

It seems like there are many Ford(12) and Chevrolet(20) involved in fatal accident. Followed by Honda(37) and Toyota(49). However, I would not accuse them for the fatality of the accident because this does not account the total number of vehicle sold. I would dive deeper into this matter and conclude if the make of the vehicle is actually "dangerous".

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

geoloc <- readxl::read_xlsx("FRPP_GLC_-_United_StatesNov42021.xlsx")

geoloc <- geoloc %>%
  select(c(`State Name`, `State Code`)) %>%
  mutate(
    State_Name = `State Name`,
    State_Code = as.integer(`State Code`)
  ) %>%
  select(c(State_Name,State_Code)) %>%
  distinct()

str(geoloc)

joined_geoloc <- make_by_fatal_hit %>%
  left_join(geoloc, by = c("STATE" = "State_Code")) %>%
  mutate(
    state_name = tolower(State_Name)
  )

library(tidyverse)
states <- map_data('state')         

stateName <- states %>% group_by(region) %>% summarize(long=mean(long), lat=mean(lat)) %>% left_join(joined_geoloc, by =c("region" = "state_name")) %>% mutate(
  statenameandmake = paste(region, MAKE, sep = "\n")
)

states %>% ggplot(aes(x = long, y = lat)) + 
  geom_polygon(aes(group = group)) +
  geom_text(aes(label=statenameandmake), color='white', data=stateName) +
  geom_path(aes(group=group))
  coord_map()

```


***4.Join the accident and person table (work out which variable(s) to use)***

ST_CASE 
should be used to merge the Person data file with the Accident data file for a set of all motorists 
and non-motorists.

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

mergedaccper <- per %>%
  left_join(acc, by = c("ST_CASE" = "ST_CASE"))

```

***5.Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results and explain what you find.***

Weekends tend to be more accidents than the weekdays.
Hour after 12PM is more dangerous than others, peaked in 6PM.
Male tend to be in more accident than female.

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

bydayweek <- mergedaccper %>%
  group_by(DAY_WEEK) %>%
  mutate(
    sum = n()
  ) %>%
  select(c(DAY_WEEK, sum)) %>%
  distinct() %>%
  arrange(DAY_WEEK)

plt1 <- bydayweek %>% 
  ggplot(aes(x=DAY_WEEK, y = sum)) + geom_bar(stat = "identity") + scale_x_continuous(breaks =seq(1,7,by=1), labels = c("Sun", "Mon", "Tue","Wed","Thu","Fri","Sat"))
show(plt1)

byhour <- mergedaccper %>%
  group_by(HOUR.x) %>%
  mutate(
    sum = n()
  ) %>%
  select(c(HOUR.x, sum)) %>%
  distinct() %>%
  arrange(HOUR.x) %>%
  filter(!HOUR.x ==99)

plt2 <- byhour %>% 
  ggplot(aes(x=HOUR.x, y = sum)) + geom_bar(stat = "identity")
show(plt2)

bysex <- mergedaccper %>%
  group_by(SEX) %>%
  mutate(
    sum = n()
  ) %>%
  select(c(SEX, sum)) %>%
  distinct() %>%
  arrange(SEX) %>%
  filter(SEX == 1 | SEX ==2)

plt3 <- bysex %>% 
  ggplot(aes(x=SEX, y = sum)) + geom_bar(stat = "identity") + scale_x_continuous(breaks =seq(1,2,by=1), labels = c("Male", "Female"))
show(plt3)
```


***6.Now plot a choropleth map of the number of deaths on a county level. Also explain what you find.***

It seems like the east side of the nation suffers more death in accidents. I believe that it is largely dependent to the population and the population density. For west side, it is mostly Seattle and California areas that suffers the fatal accidents.

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

geoloc <- readxl::read_xlsx("FRPP_GLC_-_United_StatesNov42021.xlsx")

countygeoloc <- geoloc %>%
  select(c(`State Name`, `State Code`, `County Name`, `County Code`)) %>%
  mutate(
    State_Name = `State Name`,
    State_Code = as.integer(`State Code`),
    County_Name = `County Name`,
    County_Code = as.integer(`County Code`)
  ) %>%
  select(c(State_Name,State_Code,County_Name,County_Code)) %>%
  distinct()

str(countygeoloc)

byfatalcounty <- mergedaccper %>%
  filter(INJ_SEV == 4) %>%
  group_by(STATE.x,COUNTY.x) %>%
  mutate(
    sum = n()
  ) %>%
  select(c(STATE.x, COUNTY.x, sum, LONGITUD,LATITUDE)) %>%
  distinct() %>%
  arrange(STATE.x,COUNTY.x)

joined_countygeoloc <- byfatalcounty %>%
  left_join(countygeoloc, by = c("STATE.x" = "State_Code", "COUNTY.x" = "County_Code")) %>%
  mutate(
    state_name = tolower(State_Name),
    county_name = tolower(County_Name),
    longitude = mean(LONGITUD),
    latitude = mean(LATITUDE)
  ) %>%
  ungroup() %>%
  mutate(
    maxpercentage = sum/max(sum)
  ) %>%
  select(c(state_name,county_name, longitude,latitude,maxpercentage,sum)) %>%
  distinct()

library(tidyverse)
states <- map_data('state')         

ggplot(states, aes(x=long, y=lat)) + 
  geom_polygon(aes(group=group)) + 
  geom_point(aes(x=longitude, y=latitude), data=joined_countygeoloc,
             color='lightgreen',
             alpha=joined_countygeoloc$maxpercentage*10, size=1) + 
  xlim(-130, -60) + ylim(20, 50) +
  coord_map()
```